---
title: "Assignment 1 - Basics in R and RMarkdown"
author: "Shefali Mhadadalkar (spm2172)"
date: "September 15, 2019"
output: 
    html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(error=TRUE,cache=TRUE)
```

# Question 1

__What are the four main types of variables? What are two functions that can be used to determine a variable type?__

The four main types of variables are __numeric__, __character__, __date__, and __logical__.

1. __numeric__ handles integers and decimals, both positive and negative along with zero.
2. __character__ handles strings, as `character` or `factor` variables.
3. __date__ handles dates and times, with `Date` and `POSIXct` as the most common types storing date and datetime respectively.
4. __logical__ handles Boolean values `TRUE` and `FALSE`.

Two of the functions that can be used to test variable data types are `class` and `typeof`.

```{r}
x <- 3.4
class(x)
```

```{r}
x <- "hello"
typeof(x)
```

# Question 2

__Create a vector of numbers from one to ten called numVec. Use the sum function to calculate the sum of this vector.__

```{r}
numVec <- 1:10
sum(numVec)

```

# Question 3

__Write a function that has two arguments. If the first argument is bigger, return the sum of the numbers. If the second argument is bigger, return the product of the numbers.__

```{r}
TestFunction <- function(arg1, arg2)
{
  if(arg1 > arg2)
  {
    print(arg1 +  arg2)
  }else if(arg1 < arg2)
  {
    print(arg1 * arg2)
  }else
  {
    print("You entered equal values!")
  }
}

TestFunction(5,1)
TestFunction(1,3)
TestFunction(0,0)
```

# Question 4

__What data type does read_csv return?__

__read_csv__ returns a `data.frame`, tightly coupled collections of variables which share many of the properties of matrices and of lists, used as the fundamental data structure by most of R's modeling software.

# Question 5

__What data type does read_excel return?__

__read_excel__ returns a `tibble` instead of a`data.frame`, a better version of data frames.

# Question 6

__Read the acs_ny.csv data from [https://jaredlander.com/data/](https://www.jaredlander.com/data). Display for the first 10 rows and the last 10 rows.__

```{r}
acsdata = read.csv(file = "https://jaredlander.com/data/acs_ny.csv", header = TRUE)

print("Top 10 rows displayed below:")
head(acsdata, 10)

print("Bottom 10 rows displayed below:")
tail(acsdata, 10)

```

